---
title: "putting data"
output: html_notebook
---

#outline

* dailyUpdate script for
    * pulling raw Google Forms data 
    * wrangle into tidy source data
* data objects
    * Google Forms raw response data
    * tidyPutts: primary source for viz/table displays
    * individualRounds:  table of individual game data


# to do
> add to tidyPutts:
    > circle2 (0/1)
    > game result (W/L/T)
> rework as needed: individualRounds, holeSummary, # hole difficulty




# dailyPuttingStatUpdate.R
```{r}

library(tidyverse)
library(googlesheets4)

## import raw scoring data from gsheet 'Disc Putting Scorecard_BH1 (Responses)'
a <- read_sheet("https://docs.google.com/spreadsheets/d/1zsM6NH5NoT_QI7be9z1VnkurByc9DsmGCzeoomI5AoY/edit#gid=184111277")
names(a) <- make.names(names(a))
gsraw <- as.data.frame(a) %>% 
  mutate(Special.Notes = as.character(Special.Notes))

circle2s <- read_sheet("https://docs.google.com/spreadsheets/d/1qmQB_rLHvEdNmKP3omnVYM2yn5UlibUx3y-WmrSZij8/edit#gid=1610389370", sheet = "circle2s")

## extract separate object for each game type & position
### tidy solo rounds
tidySolo <- gsraw %>%
  filter(grepl("solo", Game.style)) %>% 
  mutate(player = if_else(PlayerID == "OTHER_WRITE_IN",
                          paste0("WRITE_IN", home.solo.WRITE.IN.nickname.number), 
                          PlayerID)) %>% 
  mutate(position = "solo") %>% 
  mutate(course=Putting.Course) %>% 
  #select(course=Putting.Course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, starts_with("solo.putts.made"))  %>% 
  pivot_longer(c("solo.putts.made..hole1.", "solo.putts.made..hole2.", 
                 "solo.putts.made..hole3.", "solo.putts.made..hole4.", "solo.putts.made..hole5.", 
                 "solo.putts.made..hole6.", "solo.putts.made..hole7.", "solo.putts.made..hole8.", 
                 "solo.putts.made..hole9.", "solo.putts.made..hole10.", "solo.putts.made..hole11.", 
                 "solo.putts.made..hole12.", "solo.putts.made..hole13.", "solo.putts.made..hole14.", 
                 "solo.putts.made..hole15.", "solo.putts.made..hole16.", "solo.putts.made..hole17.", 
                 "solo.putts.made..hole18."), names_to = "hole", values_to = "result") %>% 
  mutate(hole = as.integer(gsub("solo.putts.made..hole", "", hole))) %>% 
  mutate(putts.made = if_else(is.na(result), 0, 
                              ifelse(result == "putt made", 1, 999))) %>%
  left_join(circle2s, by = c("course", "hole")) %>% 
  mutate(circle2 = if_else(is.na(circle2), 0, 1)) %>% 
  mutate(score = if_else(putts.made == 0, 0, 
                         ifelse(putts.made == 1, 1+circle2, 999))) %>% 
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made, score, circle2)


### tidy 1-on-1 (h2h) rounds part1:  home player
tidy1on1_home <- gsraw %>% filter(Game.style == "1-on-1") %>% 
  mutate(player = if_else(Home == "OTHER_WRITE_IN",
                          paste0("WRITE_IN", home.solo.WRITE.IN.nickname.number), 
                          Home)) %>% 
  mutate(position = "home") %>% 
  mutate(course=Putting.Course) %>% 
  #select(starts_with("h2h."))  %>% 
  pivot_longer(c( "h2h.putts.made..hole1.", 
                  "h2h.putts.made..hole2.", "h2h.putts.made..hole3.", "h2h.putts.made..hole4.", 
                  "h2h.putts.made..hole5.", "h2h.putts.made..hole6.", "h2h.putts.made..hole7.", 
                  "h2h.putts.made..hole8.", "h2h.putts.made..hole9.", "h2h.putts.made..hole10.", 
                  "h2h.putts.made..hole11.", "h2h.putts.made..hole12.", "h2h.putts.made..hole13.", 
                  "h2h.putts.made..hole14.", "h2h.putts.made..hole15.", "h2h.putts.made..hole16.", 
                  "h2h.putts.made..hole17.", "h2h.putts.made..hole18."), names_to = "hole", values_to = "result") %>% 
  mutate(hole = as.integer(gsub("h2h.putts.made..hole", "", hole))) %>% 
  mutate(putts.made = if_else(is.na(result), 0, 
                              ifelse(grepl("home", result)==TRUE, 1, 0))) %>%
  left_join(circle2s, by = c("course", "hole")) %>% 
  mutate(circle2 = if_else(is.na(circle2), 0, 1)) %>% 
  mutate(score = if_else(putts.made == 0, 0, 
                         ifelse(putts.made == 1, 1+circle2, 999))) %>% 
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made, score, circle2)

### tidy 1-on-1 (h2h) rounds part2:  visiting player
tidy1on1_visitor <- gsraw %>% filter(Game.style == "1-on-1") %>% 
  mutate(player = if_else(Visitor == "OTHER_WRITE_IN",
                          paste0("WRITE_IN", visitor.player2.WRITE.IN.nickname.number), 
                          Visitor)) %>% 
  mutate(position = "visitor") %>% 
  mutate(course=Putting.Course) %>% 
  #select(starts_with("h2h."))  %>% 
  pivot_longer(c( "h2h.putts.made..hole1.", 
                  "h2h.putts.made..hole2.", "h2h.putts.made..hole3.", "h2h.putts.made..hole4.", 
                  "h2h.putts.made..hole5.", "h2h.putts.made..hole6.", "h2h.putts.made..hole7.", 
                  "h2h.putts.made..hole8.", "h2h.putts.made..hole9.", "h2h.putts.made..hole10.", 
                  "h2h.putts.made..hole11.", "h2h.putts.made..hole12.", "h2h.putts.made..hole13.", 
                  "h2h.putts.made..hole14.", "h2h.putts.made..hole15.", "h2h.putts.made..hole16.", 
                  "h2h.putts.made..hole17.", "h2h.putts.made..hole18."), names_to = "hole", values_to = "result") %>% 
  mutate(hole = as.integer(gsub("h2h.putts.made..hole", "", hole))) %>% 
  mutate(putts.made = if_else(is.na(result), 0, 
                              ifelse(grepl("visitor", result)==TRUE, 1, 0))) %>%
  left_join(circle2s, by = c("course", "hole")) %>% 
  mutate(circle2 = if_else(is.na(circle2), 0, 1)) %>% 
  mutate(score = if_else(putts.made == 0, 0, 
                         ifelse(putts.made == 1, 1+circle2, 999))) %>% 
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made, score, circle2)


### tidy 3way rounds rounds part1 of 3:  home player (aka player1)
tidy3way_player1 <- gsraw %>% filter(Game.style == "3way") %>% 
  mutate(player = if_else(player1 == "OTHER_WRITE_IN",
                          paste0("WRITE_IN", home.solo.WRITE.IN.nickname.number), 
                          player1)) %>% 
  mutate(position = "player1") %>% 
  mutate(course=Putting.Course) %>% 
  #select(starts_with("X3way."))  %>% 
  pivot_longer(c("X3way.putts.made..hole1.", "X3way.putts.made..hole2.", "X3way.putts.made..hole3.", 
                 "X3way.putts.made..hole4.", "X3way.putts.made..hole5.", "X3way.putts.made..hole6.", 
                 "X3way.putts.made..hole7.", "X3way.putts.made..hole8.", "X3way.putts.made..hole9.", 
                 "X3way.putts.made..hole10.", "X3way.putts.made..hole11.", "X3way.putts.made..hole12.", 
                 "X3way.putts.made..hole13.", "X3way.putts.made..hole14.", "X3way.putts.made..hole15.", 
                 "X3way.putts.made..hole16.", "X3way.putts.made..hole17.", "X3way.putts.made..hole18."), names_to = "hole", values_to = "result") %>% 
  mutate(hole = as.integer(gsub("X3way.putts.made..hole", "", hole))) %>% 
  mutate(putts.made = if_else(is.na(result), 0, 
                              ifelse(grepl("player1", result)==TRUE, 1, 0))) %>%
  left_join(circle2s, by = c("course", "hole")) %>% 
  mutate(circle2 = if_else(is.na(circle2), 0, 1)) %>% 
  mutate(score = if_else(putts.made == 0, 0, 
                         ifelse(putts.made == 1, 1+circle2, 999))) %>% 
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made, score, circle2)

### tidy 3way rounds rounds part2 of3:  player2
tidy3way_player2 <- gsraw %>% filter(Game.style == "3way") %>% 
  mutate(player = if_else(player2 == "OTHER_WRITE_IN",
                          paste0("WRITE_IN", visitor.player2.WRITE.IN.nickname.number), 
                          player2)) %>% 
  mutate(position = "player2") %>% 
  mutate(course=Putting.Course) %>% 
  #select(starts_with("X3way."))  %>% 
  pivot_longer(c("X3way.putts.made..hole1.", "X3way.putts.made..hole2.", "X3way.putts.made..hole3.", 
                 "X3way.putts.made..hole4.", "X3way.putts.made..hole5.", "X3way.putts.made..hole6.", 
                 "X3way.putts.made..hole7.", "X3way.putts.made..hole8.", "X3way.putts.made..hole9.", 
                 "X3way.putts.made..hole10.", "X3way.putts.made..hole11.", "X3way.putts.made..hole12.", 
                 "X3way.putts.made..hole13.", "X3way.putts.made..hole14.", "X3way.putts.made..hole15.", 
                 "X3way.putts.made..hole16.", "X3way.putts.made..hole17.", "X3way.putts.made..hole18."), names_to = "hole", values_to = "result") %>% 
  mutate(hole = as.integer(gsub("X3way.putts.made..hole", "", hole))) %>% 
  mutate(putts.made = if_else(is.na(result), 0, 
                              ifelse(grepl("player2", result)==TRUE, 1, 0))) %>%
  left_join(circle2s, by = c("course", "hole")) %>% 
  mutate(circle2 = if_else(is.na(circle2), 0, 1)) %>% 
  mutate(score = if_else(putts.made == 0, 0, 
                         ifelse(putts.made == 1, 1+circle2, 999))) %>% 
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made, score, circle2)

### tidy 3way rounds rounds part3 of3:  player3
tidy3way_player3 <- gsraw %>% filter(Game.style == "3way") %>% 
  mutate(player = if_else(player3 == "OTHER_WRITE_IN",
                          paste0("WRITE_IN", player3.WRITE.IN.nickname.number), 
                          player3)) %>% 
  mutate(position = "player3") %>% 
  mutate(course=Putting.Course) %>% 
  #select(starts_with("X3way."))  %>% 
  pivot_longer(c("X3way.putts.made..hole1.", "X3way.putts.made..hole2.", "X3way.putts.made..hole3.", 
                 "X3way.putts.made..hole4.", "X3way.putts.made..hole5.", "X3way.putts.made..hole6.", 
                 "X3way.putts.made..hole7.", "X3way.putts.made..hole8.", "X3way.putts.made..hole9.", 
                 "X3way.putts.made..hole10.", "X3way.putts.made..hole11.", "X3way.putts.made..hole12.", 
                 "X3way.putts.made..hole13.", "X3way.putts.made..hole14.", "X3way.putts.made..hole15.", 
                 "X3way.putts.made..hole16.", "X3way.putts.made..hole17.", "X3way.putts.made..hole18."), names_to = "hole", values_to = "result") %>% 
  mutate(hole = as.integer(gsub("X3way.putts.made..hole", "", hole))) %>% 
  mutate(putts.made = if_else(is.na(result), 0, 
                              ifelse(grepl("player3", result)==TRUE, 1, 0))) %>%
  left_join(circle2s, by = c("course", "hole")) %>% 
  mutate(circle2 = if_else(is.na(circle2), 0, 1)) %>% 
  mutate(score = if_else(putts.made == 0, 0, 
                         ifelse(putts.made == 1, 1+circle2, 999))) %>% 
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made, score, circle2)

## rbind
tidyPutts <- rbind(tidy1on1_home, tidy1on1_visitor) %>% 
  rbind(., tidy3way_player1) %>% 
  rbind(., tidy3way_player2) %>% 
  rbind(., tidy3way_player3) %>% 
  rbind(., tidySolo)  %>% 
  arrange(desc(Timestamp))



## Individual Rounds
# calculate winner
WLT <- tidyPutts %>% group_by(Game.style, Timestamp, player) %>% 
  summarise_at(.,vars(score), 
  funs(sum)) %>% 
  mutate(WLT = if_else(Game.style == "solo", "solo", 
                       #if_else(Game.style == "3way", "3way", 
                               if_else(score < max(score), "L",
           if_else(score > min(score) & score > median(score), "W",
                       "T")))) %>% 
  ungroup() %>% 
   group_by(player, WLT) %>% tally() %>% pivot_wider(names_from = WLT, values_from = n) %>% arrange(desc(W, T)) %>% 
  select(player, W, L, T, solo)

WLT[is.na(WLT)] <- 0 ## replace NA with zero

WLT <- WLT %>% 
  mutate(points = 2*W +T) %>% 
  mutate(winPct = round(W/(W+L+T), 3)) %>% 
  arrange(desc(points))

individualRounds <- tidyPutts %>% 
  mutate(chartime = as.character(Timestamp)) %>% 
  group_by(course, player, Timestamp, Game.style, position, Purpose, Outcome) %>% 
  summarise_at(.,vars(putts.made, score), 
               funs(sum(., na.rm = TRUE))) %>% 
  ungroup() %>% 
  mutate(circle2s = score-putts.made) %>% 
  arrange(desc(score)) %>% 
  select(player, score, circle2s, putts.made, everything())







holeSummary <- tidyPutts %>% 
  group_by(course, hole) %>% 
  summarise_at(.,vars(putts.made), 
               funs(n(), sum(., na.rm = TRUE))) %>%
  mutate(attempts = n, made = sum) %>%
  mutate(madePct =round(100*(made/attempts), 1)) %>% 
  select(course, hole, attempts, made, madePct)




## league average
courseMeanScore <- individualRounds %>%
  group_by(course) %>%
  summarise_at(.,vars(score, putts.made), 
               funs(n(), mean(., na.rm = TRUE), sd(., na.rm = TRUE))) %>% 
  mutate(leagueMean = score_mean) %>%
  mutate(leagueSD = score_sd) %>%
  select(course, leagueMean, leagueSD)

##player summary
playerCourseSummary <- individualRounds %>% 
  left_join(courseMeanScore, by = "course") %>% 
  mutate(gWAR = leagueSD*(score - leagueMean)) %>% 
  group_by(player, course) %>% 
  summarise_at(.,vars(score, putts.made, circle2s), 
               funs(sd(., na.rm = TRUE),mean(., na.rm = TRUE), min(., na.rm = TRUE), max(., na.rm = TRUE), median(., na.rm = TRUE), sum(., na.rm = TRUE), n())) %>% 
  select(player, course, rounds=score_n, score_mean, circle2s_sum, putts.made_sum) %>% 
  arrange(desc(score_mean))



##putting WAR
puttingWAR <- individualRounds %>% 
  left_join(courseMeanScore, by = "course") %>% 
  mutate(gWAR = leagueSD*(score - leagueMean)) %>% 
  group_by(player, course) %>% 
  summarise_at(.,vars(gWAR), 
               funs(mean(., na.rm = TRUE), n())) %>% 
  arrange(desc(mean))



data.frame(x=Sys.time()) %>%  ## GMT +4?
  mutate(StatsLastUpdated = format(x, tz="America/New_York")) %>% 
  select(-x) %>% 
write_sheet(., ss = "https://docs.google.com/spreadsheets/d/13T0Y1pAxqhL7rzRcFwtnFZPUFHeBIldcqLk1RF172nM/edit#gid=1215521890", sheet = "StatsLastUpdated")

#write local copies
gsraw %>%  
  mutate(Special.Notes = as.character(Special.Notes)) %>% 
  write.csv(paste0("C:\\!RlocalProjectsGit\\putting-leaderboards-1\\Disc Putting Form Responses 1_", strftime(Sys.time(), "%Y%m%d%H%M"),  ".csv"))

individualRounds %>%  
  write.csv(paste0("C:\\!RlocalProjectsGit\\putting-leaderboards-1\\Disc Putting Rounds_", strftime(Sys.time(), "%Y%m%d%H%M"),  ".csv"))

# write tables to googlesheet
write_sheet(WLT, ss = "https://docs.google.com/spreadsheets/d/13T0Y1pAxqhL7rzRcFwtnFZPUFHeBIldcqLk1RF172nM/edit#gid=878454155", sheet = "WLT")

write_sheet(individualRounds, ss = "https://docs.google.com/spreadsheets/d/13T0Y1pAxqhL7rzRcFwtnFZPUFHeBIldcqLk1RF172nM/edit#gid=878454155", sheet = "individualRounds")

write_sheet(holeSummary, ss = "https://docs.google.com/spreadsheets/d/13T0Y1pAxqhL7rzRcFwtnFZPUFHeBIldcqLk1RF172nM/edit#gid=878454155", sheet = "holeSummary")

write_sheet(playerCourseSummary, ss = "https://docs.google.com/spreadsheets/d/13T0Y1pAxqhL7rzRcFwtnFZPUFHeBIldcqLk1RF172nM/edit#gid=878454155", sheet = "playerCourseSummary")

write_sheet(puttingWAR, ss = "https://docs.google.com/spreadsheets/d/13T0Y1pAxqhL7rzRcFwtnFZPUFHeBIldcqLk1RF172nM/edit#gid=878454155", sheet = "puttingWAR")


```


# top10 rounds (all-time, according to the data)
```{r}
a <- individualRounds %>% 
  select(course, player, putts.made, circle2s, score, Game.style, Purpose, Outcome, Timestamp) %>% 
  arrange(desc(score)) %>% 
  head(10)

DT::datatable(a, filter = 'top')
```


# course difficulty
```{r}
a <- individualRounds %>%
  group_by(course) %>%
  summarise_at(.,vars(score, putts.made), 
               funs(n(), mean(., na.rm = TRUE), median(., na.rm = TRUE), min(., na.rm = TRUE), max(., na.rm = TRUE))) %>% 
  mutate(totalRounds = score_n) %>% 
  select(-score_n, -putts.made_n)

b <- data.frame(t(a))
names(b) <- b[1,]
c <- b[-1,] %>% rownames_to_column("statistic") %>% arrange((statistic))

DT::datatable(c)
```


# hole difficulty
```{r}

a <- tidyPutts %>% 
  filter(course == "Beggars' Knob") %>% 
  mutate(Special.Notes = as.character(Special.Notes)) %>% 
  mutate(hole = if_else(hole<10, paste0("hole0", hole), paste0("hole", hole))) %>% 
  group_by(player, course, hole) %>% 
  summarise_at(.,vars(putts.made), 
  funs(mean(., na.rm = TRUE), n())) %>% 
  filter(n >1) %>% 
  mutate(madePct = round(100*mean, 1))


ggplot(a, aes(hole,madePct, fill = course)) + geom_boxplot() + stat_summary(fun.y=mean, geom="point", shape=5, size=4) + labs(subtitle="a", y="madePct", x="hole", title="Boxplots of player averages for each hole", caption= paste0("summary as of ", as.character(Sys.time()))) + theme(aspect.ratio = 0.33) + theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1))

 a %>% 
   select(course, player, hole, madePct, n) %>% 
  DT::datatable(., filter = 'top')

```


# Datatable:  all putts ever
```{r}
tidyPutts %>% 
  mutate(Special.Notes = as.character(Special.Notes)) %>% 
  select(player, course, hole, putts.made, everything()) %>% 
  DT::datatable(., filter = 'top')

```





## Team scores
```{r}
### TEAM rounds
df <- gsraw %>% filter(Game.style == "2-on-2")
```

