---
title: "Disc Putting Championship Homepage"
output: 
  html_document:
    code_folding: hide
    toc: true
    theme: united
    toc_float: true
    df_print: paged
---

$~$

$~$

$~$

* Why putt discs?
    * Disc golf is one of the fastest-growing and most accessible outdoor activities in the US and many other countries. 
    * To putt well is to play well. 
    * To play well is to live well.

* Why collect or analyze disc putting data?
    * This is a demonstration of reproducible research and transparency in the process of tabulating statistics from raw game data. Each approved putting course is a set of 18 well-defined shots, typically comprised of three circle2 putts and 15 circle1 putts. With unlimited numbers of players putting the exact same shots, patterns emerge that provide unique insights into disc putting.  


$~$

$~$

$~$

# Player stats
```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(googlesheets4)
## import raw scoring data from gsheet 'Disc Putting Scorecard_BH1 (Responses)'
a <- read_sheet("https://docs.google.com/spreadsheets/d/1zsM6NH5NoT_QI7be9z1VnkurByc9DsmGCzeoomI5AoY/edit#gid=184111277")
names(a) <- make.names(names(a))
gsraw <- as.data.frame(a) %>% 
  mutate(Special.Notes = as.character(Special.Notes))
circle2s <- read_sheet("https://docs.google.com/spreadsheets/d/1qmQB_rLHvEdNmKP3omnVYM2yn5UlibUx3y-WmrSZij8/edit#gid=1610389370", sheet = "circle2s")
## extract separate object for each game type & position
### tidy solo rounds
tidySolo <- gsraw %>%
  filter(grepl("solo", Game.style)) %>% 
  mutate(player = if_else(PlayerID == "OTHER_WRITE_IN",
                          paste0("WRITE_IN", home.solo.WRITE.IN.nickname.number), 
                          PlayerID)) %>% 
  mutate(position = "solo") %>% 
  mutate(course=Putting.Course) %>% 
  #select(course=Putting.Course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, starts_with("solo.putts.made"))  %>% 
  pivot_longer(c("solo.putts.made..hole1.", "solo.putts.made..hole2.", 
                 "solo.putts.made..hole3.", "solo.putts.made..hole4.", "solo.putts.made..hole5.", 
                 "solo.putts.made..hole6.", "solo.putts.made..hole7.", "solo.putts.made..hole8.", 
                 "solo.putts.made..hole9.", "solo.putts.made..hole10.", "solo.putts.made..hole11.", 
                 "solo.putts.made..hole12.", "solo.putts.made..hole13.", "solo.putts.made..hole14.", 
                 "solo.putts.made..hole15.", "solo.putts.made..hole16.", "solo.putts.made..hole17.", 
                 "solo.putts.made..hole18."), names_to = "hole", values_to = "result") %>% 
  mutate(hole = as.integer(gsub("solo.putts.made..hole", "", hole))) %>% 
  mutate(putts.made = if_else(is.na(result), 0, 
                              ifelse(result == "putt made", 1, 999)))  %>%
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made)
### tidy 1-on-1 (h2h) rounds part1:  home player
tidy1on1_home <- gsraw %>% filter(Game.style == "1-on-1") %>% 
  mutate(player = if_else(Home == "OTHER_WRITE_IN",
                          paste0("WRITE_IN", home.solo.WRITE.IN.nickname.number), 
                          Home)) %>% 
  mutate(position = "home") %>% 
  mutate(course=Putting.Course) %>% 
  #select(starts_with("h2h."))  %>% 
  pivot_longer(c( "h2h.putts.made..hole1.", 
                  "h2h.putts.made..hole2.", "h2h.putts.made..hole3.", "h2h.putts.made..hole4.", 
                  "h2h.putts.made..hole5.", "h2h.putts.made..hole6.", "h2h.putts.made..hole7.", 
                  "h2h.putts.made..hole8.", "h2h.putts.made..hole9.", "h2h.putts.made..hole10.", 
                  "h2h.putts.made..hole11.", "h2h.putts.made..hole12.", "h2h.putts.made..hole13.", 
                  "h2h.putts.made..hole14.", "h2h.putts.made..hole15.", "h2h.putts.made..hole16.", 
                  "h2h.putts.made..hole17.", "h2h.putts.made..hole18."), names_to = "hole", values_to = "result") %>% 
  mutate(hole = as.integer(gsub("h2h.putts.made..hole", "", hole))) %>% 
  mutate(putts.made = if_else(is.na(result), 0, 
                              ifelse(grepl("home", result)==TRUE, 1, 0)))  %>%
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made)
### tidy 1-on-1 (h2h) rounds part2:  visiting player
tidy1on1_visitor <- gsraw %>% filter(Game.style == "1-on-1") %>% 
  mutate(player = if_else(Visitor == "OTHER_WRITE_IN",
                          paste0("WRITE_IN", visitor.player2.WRITE.IN.nickname.number), 
                          Visitor)) %>% 
  mutate(position = "visitor") %>% 
  mutate(course=Putting.Course) %>% 
  #select(starts_with("h2h."))  %>% 
  pivot_longer(c( "h2h.putts.made..hole1.", 
                  "h2h.putts.made..hole2.", "h2h.putts.made..hole3.", "h2h.putts.made..hole4.", 
                  "h2h.putts.made..hole5.", "h2h.putts.made..hole6.", "h2h.putts.made..hole7.", 
                  "h2h.putts.made..hole8.", "h2h.putts.made..hole9.", "h2h.putts.made..hole10.", 
                  "h2h.putts.made..hole11.", "h2h.putts.made..hole12.", "h2h.putts.made..hole13.", 
                  "h2h.putts.made..hole14.", "h2h.putts.made..hole15.", "h2h.putts.made..hole16.", 
                  "h2h.putts.made..hole17.", "h2h.putts.made..hole18."), names_to = "hole", values_to = "result") %>% 
  mutate(hole = as.integer(gsub("h2h.putts.made..hole", "", hole))) %>% 
  mutate(putts.made = if_else(is.na(result), 0, 
                              ifelse(grepl("visitor", result)==TRUE, 1, 0)))  %>%
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made)
### tidy 3way rounds rounds part1 of 3:  home player (aka player1)
tidy3way_player1 <- gsraw %>% filter(Game.style == "3way") %>% 
  mutate(player = if_else(player1 == "OTHER_WRITE_IN",
                          paste0("WRITE_IN", home.solo.WRITE.IN.nickname.number), 
                          player1)) %>% 
  mutate(position = "player1") %>% 
  mutate(course=Putting.Course) %>% 
  #select(starts_with("X3way."))  %>% 
  pivot_longer(c("X3way.putts.made..hole1.", "X3way.putts.made..hole2.", "X3way.putts.made..hole3.", 
                 "X3way.putts.made..hole4.", "X3way.putts.made..hole5.", "X3way.putts.made..hole6.", 
                 "X3way.putts.made..hole7.", "X3way.putts.made..hole8.", "X3way.putts.made..hole9.", 
                 "X3way.putts.made..hole10.", "X3way.putts.made..hole11.", "X3way.putts.made..hole12.", 
                 "X3way.putts.made..hole13.", "X3way.putts.made..hole14.", "X3way.putts.made..hole15.", 
                 "X3way.putts.made..hole16.", "X3way.putts.made..hole17.", "X3way.putts.made..hole18."), names_to = "hole", values_to = "result") %>% 
  mutate(hole = as.integer(gsub("X3way.putts.made..hole", "", hole))) %>% 
  mutate(putts.made = if_else(is.na(result), 0, 
                              ifelse(grepl("player1", result)==TRUE, 1, 0))) %>%
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made)
### tidy 3way rounds rounds part2 of3:  player2
tidy3way_player2 <- gsraw %>% filter(Game.style == "3way") %>% 
  mutate(player = if_else(player2 == "OTHER_WRITE_IN",
                          paste0("WRITE_IN", visitor.player2.WRITE.IN.nickname.number), 
                          player2)) %>% 
  mutate(position = "player2") %>% 
  mutate(course=Putting.Course) %>% 
  #select(starts_with("X3way."))  %>% 
  pivot_longer(c("X3way.putts.made..hole1.", "X3way.putts.made..hole2.", "X3way.putts.made..hole3.", 
                 "X3way.putts.made..hole4.", "X3way.putts.made..hole5.", "X3way.putts.made..hole6.", 
                 "X3way.putts.made..hole7.", "X3way.putts.made..hole8.", "X3way.putts.made..hole9.", 
                 "X3way.putts.made..hole10.", "X3way.putts.made..hole11.", "X3way.putts.made..hole12.", 
                 "X3way.putts.made..hole13.", "X3way.putts.made..hole14.", "X3way.putts.made..hole15.", 
                 "X3way.putts.made..hole16.", "X3way.putts.made..hole17.", "X3way.putts.made..hole18."), names_to = "hole", values_to = "result") %>% 
  mutate(hole = as.integer(gsub("X3way.putts.made..hole", "", hole))) %>% 
  mutate(putts.made = if_else(is.na(result), 0, 
                              ifelse(grepl("player2", result)==TRUE, 1, 0)))  %>%
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made)
### tidy 3way rounds rounds part3 of3:  player3
tidy3way_player3 <- gsraw %>% filter(Game.style == "3way") %>% 
  mutate(player = if_else(player3 == "OTHER_WRITE_IN",
                          paste0("WRITE_IN", player3.WRITE.IN.nickname.number), 
                          player3)) %>% 
  mutate(position = "player3") %>% 
  mutate(course=Putting.Course) %>% 
  #select(starts_with("X3way."))  %>% 
  pivot_longer(c("X3way.putts.made..hole1.", "X3way.putts.made..hole2.", "X3way.putts.made..hole3.", 
                 "X3way.putts.made..hole4.", "X3way.putts.made..hole5.", "X3way.putts.made..hole6.", 
                 "X3way.putts.made..hole7.", "X3way.putts.made..hole8.", "X3way.putts.made..hole9.", 
                 "X3way.putts.made..hole10.", "X3way.putts.made..hole11.", "X3way.putts.made..hole12.", 
                 "X3way.putts.made..hole13.", "X3way.putts.made..hole14.", "X3way.putts.made..hole15.", 
                 "X3way.putts.made..hole16.", "X3way.putts.made..hole17.", "X3way.putts.made..hole18."), names_to = "hole", values_to = "result") %>% 
  mutate(hole = as.integer(gsub("X3way.putts.made..hole", "", hole))) %>% 
  mutate(putts.made = if_else(is.na(result), 0, 
                              ifelse(grepl("player3", result)==TRUE, 1, 0))) %>%
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made)
## rbind and add course info
tidyPutts <- rbind(tidy1on1_home, tidy1on1_visitor) %>% 
  rbind(., tidy3way_player1) %>% 
  rbind(., tidy3way_player2) %>% 
  rbind(., tidy3way_player3) %>% 
  rbind(., tidySolo)  %>% 
  left_join(circle2s, by = c("course", "hole")) %>% 
  mutate(Special.Notes = as.character(Special.Notes)) %>% 
  mutate(circle2 = if_else(is.na(circle2), 0, 1)) %>% 
  mutate(score = if_else(putts.made == 0, 0, 
                         ifelse(putts.made == 1, 1+circle2, 999))) %>% 
  arrange(desc(Timestamp)) %>%
  
  select(course, Timestamp, Game.style, position, Purpose, Outcome, Special.Notes, player, hole, putts.made, score, circle2)
## Individual Rounds
# calculate winner
WLT <- tidyPutts %>% 
  group_by(Game.style, Timestamp, player) %>% 
  summarise_at(.,vars(score), 
               funs(sum)) %>% 
  mutate(WLT = if_else(Game.style == "solo", "solo", 
                       #if_else(Game.style == "3way", "3way", 
                       if_else(score < max(score), "L",
                               if_else(score > min(score) & score > median(score), "W",
                                       "T")))) %>% 
  ungroup() %>% 
  group_by(player, WLT) %>% tally() %>% 
  ungroup() %>% 
  pivot_wider(names_from = WLT, values_from = n) %>% arrange(desc(W, T)) %>% 
  select(player, W, L, T)
WLT[is.na(WLT)] <- 0 ## replace NA with zero
WLT <- WLT %>% 
  mutate(points = 2*W +T) %>% 
  mutate(winPct = round(W/(W+L+T), 3)) %>% 
  mutate(winPct = format(round(winPct, digits = 3), nsmall = 3)) %>%
  select(player, winPct, everything()) %>% 
  arrange(desc(points))
individualRounds <- tidyPutts %>% 
  mutate(chartime = as.character(Timestamp)) %>% 
  group_by(course, player, Timestamp, Game.style, position, Purpose, Outcome) %>% 
  summarise_at(.,vars(putts.made, score), 
               funs(sum(., na.rm = TRUE))) %>% 
  ungroup() %>% 
  mutate(circle2s = score-putts.made) %>% 
  arrange(desc(score)) %>% 
  select(player, score, circle2s, putts.made, everything())
courseHoleSummary <- tidyPutts %>% 
  group_by(course, hole) %>% 
  summarise_at(.,vars(putts.made), 
               funs(n(), sum(., na.rm = TRUE))) %>%
  ungroup() %>% 
  mutate(attempts = n, made = sum) %>%
  mutate(madePct =round(100*(made/attempts), 1)) %>% 
  mutate(holeChar = paste0("hole", hole)) %>% 
  select(course, hole, attempts, made, madePct)
## Average score for each course >> courseMeanScore
courseMeanScore <- individualRounds %>%
  group_by(course) %>%
  summarise_at(.,vars(score, putts.made), 
               funs(n(), mean(., na.rm = TRUE), sd(., na.rm = TRUE))) %>% 
  ungroup() %>% 
  mutate(leagueMean = score_mean) %>%
  mutate(leagueSD = score_sd) %>%
  select(course, leagueMean, leagueSD)
##putting WAR
puttingWAR <- individualRounds %>% 
  left_join(courseMeanScore, by = "course") %>% 
  mutate(zscore = leagueSD*(score - leagueMean)) %>% 
  group_by(player) %>% 
  summarise_at(.,vars(zscore), 
               funs(sum(., na.rm = TRUE), n())) %>% 
  ungroup() %>% 
  mutate(x = sum/10) %>% 
  arrange(desc(x)) %>% 
  mutate(x = format(round(x, digits = 1), nsmall = 1)) %>% 
  select(player, puttingWAR = x)
## player circle1s
player_circle1 <- tidyPutts %>% filter(circle2 == 0) %>%
  group_by(player) %>% 
  summarise_at(.,vars(score, putts.made), 
               funs(mean(., na.rm = TRUE), sum(., na.rm = TRUE), n())) %>% 
  ungroup() %>% 
  mutate(x = putts.made_mean) %>% 
  arrange(desc(x), putts.made_n) %>% 
  mutate(x = format(round(x, digits = 3), nsmall = 3)) %>% 
  mutate(circle1mean = x,
         circle1attempts = putts.made_n,
         circle1made = putts.made_sum) %>% 
  select(player, circle1mean, circle1made, circle1attempts)
## player circle2s
player_circle2 <- tidyPutts %>% filter(circle2 == 1) %>%
  group_by(player) %>% 
  summarise_at(.,vars(score, putts.made), 
               funs(mean(., na.rm = TRUE), sum(., na.rm = TRUE), n())) %>% 
  ungroup() %>% 
  mutate(x = putts.made_mean) %>% 
  arrange(desc(x), putts.made_n) %>% 
  mutate(x = format(round(x, digits = 3), nsmall = 3))  %>% 
  mutate(circle2mean = x,
         circle2attempts = putts.made_n,
         circle2made = putts.made_sum) %>% 
  select(player, circle2mean, circle2made, circle2attempts)
## player mean score
player_meanScore <- individualRounds %>% 
  group_by(player) %>% 
  summarise_at(.,vars(score, putts.made), 
               funs(mean(., na.rm = TRUE), n())) %>% 
  ungroup() %>% 
  arrange(desc(score_mean)) %>% 
  mutate(score_mean = format(round(score_mean, digits = 2), nsmall = 2)) %>%
  select(player, rounds=score_n, score_mean)
## player games
player_games <- individualRounds %>% group_by(player, Game.style) %>% tally() %>% ungroup() %>% 
  pivot_wider(names_from = Game.style, values_from = n) %>% 
  select(player, "rounds 1-on-1" = "1-on-1" , "rounds 3way" = "3way", "rounds solo" = solo)
player_games[is.na(player_games)] <- 0 ## replace NA with zero
##player stats
playerStats <- WLT %>% 
  left_join(., player_meanScore, by = "player") %>% 
  left_join(., player_circle1, by = "player") %>% 
  left_join(., player_circle2, by = "player") %>% 
  left_join(., puttingWAR, by = "player")  %>% 
  left_join(., player_games, by = "player")
```


### Winning percentage*
```{r}
a <- WLT %>% 
  filter(W+L+T > 2) %>%
  arrange(desc(winPct)) %>% 
  select(Player = player, winPct, W, L, T) %>% head(10) #%>% DT::datatable(., filter = 'top') 
knitr::kable(a, align=c('l', 'c', 'c', 'c', 'c'))
```
> *minimum 3 games played, includes both 1-on-1 and 3way games.
$~$

$~$

### Average score*

```{r}
###  Average Score
a <- player_meanScore  %>% 
  filter(rounds >2) %>%
  arrange(desc(score_mean), rounds) %>% 
  head(10) %>% 
    select(Player = player, 'Average score'= score_mean, Rounds = rounds) 
knitr::kable(a, align=c('l', 'c', 'c', 'c'))
```
> *minimum 3 rounds
$~$

$~$

### Circle1 leaders*
```{r}
a <- player_circle1 %>% 
  filter(circle1attempts >9) %>% 
  head(10) %>% 
  select(Player = player, Circle1pct= circle1mean, Attempts=circle1attempts, Made = circle1made) 
knitr::kable(a, align=c('l', 'c', 'c', 'c'))
```
> *minimum 10 attempts
$~$

$~$


### Circle2 leaders*

```{r}
a <- player_circle2 %>% 
   filter(circle2attempts >2)  %>%  
  head(10) %>% 
  select(Player = player, Circle2pct= circle2mean, Attempts = circle2attempts, Made = circle2made)
knitr::kable(a, align=c('l', 'c', 'c', 'c'))
```
> *minimum 3 attempts

$~$

$~$

### Top shots*
Top 10 averages among all players on all holes.

```{r}
playerHoleSummary <- tidyPutts %>%
  group_by(course, hole, player) %>%
  summarise_at(.,vars(putts.made, score), 
               funs(n(), mean(., na.rm = TRUE), median(., na.rm = TRUE), min(., na.rm = TRUE), max(., na.rm = TRUE), sum(., na.rm = TRUE))) %>% 
  mutate(x = putts.made_mean) %>% 
  mutate(x = format(round(x, digits = 3), nsmall = 3)) %>%
  arrange(desc(x)) %>% 
  mutate(Hole = paste0(course, " - hole ", hole)) %>% 
  ungroup() %>% 
  select(Player = player, puttPct = x, Hole, Attempts = putts.made_n,  Made = putts.made_sum) 
a <- playerHoleSummary %>% 
  filter(Attempts > 9) %>%  head(10)
  
knitr::kable(a, align=c('l', 'c', 'c', 'c'))
```

> *minimum 10 attempts
$~$

$~$


### Sortable stats
```{r}
playerStats %>% DT::datatable(., filter = 'top', 
          rownames = FALSE,
          options = list(
            columnDefs = list(list(className = 'dt-center', targets = 1:17)))) ## note that column numbers start from 0, not from 1.
```



$~$

$~$

$~$

$~$

# Course stats

## Course overview
```{r}
a <- individualRounds %>%
  group_by(course) %>%
  summarise_at(.,vars(score, putts.made), 
               funs(n(), mean(., na.rm = TRUE), median(., na.rm = TRUE), min(., na.rm = TRUE), max(., na.rm = TRUE))) %>% 
  mutate(totalRounds = score_n) %>% 
  mutate(x = score_mean) %>% 
  mutate(x = format(round(x, digits = 2), nsmall = 2)) %>%
  select(course, 'Average Score' = x, 'High Score' = score_max, 'Most Putts Made' =putts.made_max, 'Total Rounds' = totalRounds)
b <- data.frame(t(a))
#b %>%  rownames_to_column("statistic")
names(b) <- b[1,]
c <- b[-1,] %>% rownames_to_column("statistic") #%>% arrange((statistic))
names(c) <- paste0("___", names(c), "___")
knitr::kable(c, align=c('r', 'c', 'c', 'c'))
```




$~$

$~$

# Contact


Disc Putting Championship - discputtingchampionship@gmail.com



$~$

$~$

# Register

#### [Player Registration](https://docs.google.com/forms/d/e/1FAIpQLSekk5AX4sNW7q8TWSEnXYJ1fe3KR8XbvAJBWFlKSCbXGHOORw/viewform)

#### [Course Registration](https://docs.google.com/forms/d/e/1FAIpQLSfcKdf9TDGmvH833hY7VjvGoKKUQCgsLKIQF-A2UbsIelfTSA/viewform)

$~$

#### [Scorecard](https://docs.google.com/forms/d/e/1FAIpQLSdZlQnmdbSv2BY1mxqzvmxMnkSR56L9buXUHy6r0ych_hm2vg/viewform)

$~$

$~$


```{r warning=FALSE, message=FALSE}
#write local copies
#gsraw %>%  mutate(Special.Notes = as.character(Special.Notes)) %>%   write.csv(paste0("C:\\!RlocalProjectsGit\\putting-leaderboards-1\\Disc Putting Form Responses 1_", strftime(Sys.time(), "%Y%m%d%H%M"),  ".csv"))
#individualRounds %>% write.csv(paste0("C:\\!RlocalProjectsGit\\putting-leaderboards-1\\individual Disc Putting Rounds_", strftime(Sys.time(), "%Y%m%d%H%M"),  ".csv"))
write_sheet(individualRounds, ss = "https://docs.google.com/spreadsheets/d/13T0Y1pAxqhL7rzRcFwtnFZPUFHeBIldcqLk1RF172nM/edit#gid=878454155", sheet = "individualRounds")
# write tables to googlesheet
write_sheet(playerStats, ss = "https://docs.google.com/spreadsheets/d/13T0Y1pAxqhL7rzRcFwtnFZPUFHeBIldcqLk1RF172nM/edit#gid=878454155", sheet = "playerStats")
write_sheet(playerHoleSummary, ss = "https://docs.google.com/spreadsheets/d/13T0Y1pAxqhL7rzRcFwtnFZPUFHeBIldcqLk1RF172nM/edit#gid=878454155", sheet = "playerHoleSummary")
write_sheet(courseHoleSummary, ss = "https://docs.google.com/spreadsheets/d/13T0Y1pAxqhL7rzRcFwtnFZPUFHeBIldcqLk1RF172nM/edit#gid=878454155", sheet = "courseHoleSummary")
#write_sheet(tidyPutts, ss = "https://docs.google.com/spreadsheets/d/13T0Y1pAxqhL7rzRcFwtnFZPUFHeBIldcqLk1RF172nM/edit#gid=878454155", sheet = "tidyPutts")
data.frame(x=Sys.time()) %>%  ## GMT +4?
  mutate(StatsLastUpdated = format(x, tz="America/New_York")) %>% 
  select(-x) %>% 
  write_sheet(., ss = "https://docs.google.com/spreadsheets/d/13T0Y1pAxqhL7rzRcFwtnFZPUFHeBIldcqLk1RF172nM/edit#gid=1215521890", sheet = "StatsLastUpdated")
mostRecentGame <- tidyPutts %>% filter(Timestamp == max(Timestamp)) %>% select(course, Timestamp) %>% unique()
```

`r paste0(mostRecentGame$Timestamp, " at ", mostRecentGame$course)` >> Most recent game in dataset

$~$

`r Sys.time()` >> Stats last updated 